Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-11 edubots_surveys.Rnw
Update edubots_surveys.Rnw
Co-authored-by: Francisco Barranco <fbarranco@ugr.es>
Reflow and some corrections in first paragraph

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers.
Apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the minds of the
surveyed. In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases, they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app-based messaging
systems vs. web-based messaging systems integrated in another learning management system).

@@

In our survey, we could not find an answer about the right moment to introducing chatbot technologies. The response to questions regarding to change of
behavior during and after the first stay-at-home stage of the pandemic do not suggest that a major (or minor, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.} or the introduction of new degrees) crisis could be an opportunity to introduce new technologies. The reason being that it does not
bring major changes in attitude 
%(as a minor ``crisis'' we could mention any institutional event that brings about methodological changes, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.}), 
%or simply the introduction of new degrees. 
Although changes driven by external forces do offer the chance of piloting new technologies, they do not
seem to bring internal changes in educators' attitudes (which, after all, is
bound to be the same). In absence of a clear answer in this direction, the right
moment to introduce new technologies is always when the IT and managing staff is
ready to support it (since, as we have seen before, ``official'' support is one of
the factors that is most valued by teachers).

@@

Initially, institutions should introduce messaging automation tools attached to
some, or several, messaging platforms that are already popular among the
university community, that would help with bureaucratic or repetitive tasks,
such as answering frequently asked questions or providing information on class
or assignment deadlines.  These chatbots will pave the way for more complex ones
that will have a greater impact on learning outcomes, such as chatbots that help
students integrate with other students in class or identify and address learning
problems in individual students or groups of them. These chatbots should also be
accompanied by analytics on student interaction and possibly some natural
language processing (in vernacular language) to help assess the general mood of
the class and how different materials or external factors affect it. These tools
would have to be attached to a learning analytics platform, which, in turn,
would be part of an institutional learning management system.

>>>>

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. The availability of certain
technologies for implementing chatbots may vary based on the outcome. In this
sense, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was any way validated in the piloting
phase of the survey, and eventually resolved by the answers.  Apparently,
anything that sends messages (even if they are not instant messages or have no
dedicated application) was the concept in the minds of the surveyed. In this
case, we would like to add a clarification. While we had in mind, as indicated
by the possible answers, instant messaging applications when we elaborated this
question, we included these institution-provided applications mainly for
completion. But it should be noted that, in most cases, they are not {\em
  instant} messaging apps in the way Telegram or Facebook Messenger are; they
are messaging facilities provided by learning management systems such as Moodle
or Blackboard. So the answer to this question must be understood in two
different ways: their preference for the {\em provider} of the technology (the
university itself, any company) as well as the {\em type} of technology
(instant, app-based messaging systems vs. web- or email-based messaging systems
integrated in another learning management system).

@@

In our survey, we could not find an answer about the right moment to introducing
chatbot technologies. The response to questions regarding to change of behavior
during and after the first stay-at-home stage of the pandemic do not suggest
that a major (or minor, such as new higher education laws\footnote{We seem to
  have one of these, at least in Spain, every 10 years or so.} or the
introduction of new degrees) crisis could be an opportunity to introduce new
technologies. The reason being that it does not bring major changes in attitude
Although changes driven by external forces do offer the chance of piloting new
technologies, they do not seem to bring internal changes in educators' attitudes
(which, after all, is bound to be the same). In absence of a clear answer in
this direction, the right moment to introduce new technologies is always when
the IT and managing staff is ready to support it (since, as we have seen before,
``official'' support is one of the factors that is most valued by teachers).

@@

Initially, institutions should introduce messaging automation tools attached to
some, or several, messaging platforms that are already popular among the
university community, that would help with bureaucratic or repetitive tasks,
such as answering frequently asked questions or providing information on class
or assignment deadlines.  These chatbots will pave the way for more complex ones
that will have a greater impact on learning outcomes, such as chatbots that help
students integrate with other students in class or identify and address learning
problems in individual students or groups of them. These chatbots should also be
accompanied by analytics on student interaction and possibly some natural
language processing (in vernacular language) to help assess the general mood of
the class and how different materials or external factors affect it. These tools
would have to be attached to a learning analytics platform, which in turn,
would be part of an institutional learning management system.
